Title of Paper

### 2017-02-17 paper.tex
minor plotting updates; update results

In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for different states will load sequential memory locations in Fig.~\ref{F:mem} in memory however the $k+1$'th species concentration will be $N$ memory locations away, increasing the likelihood of cache-misses on the CPU.
In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k\dots k+V$ (for a vector width of $V$) will access sequential memory locations if the data is stored in ``C''-order.
Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s \ll N$ (in most cases), greatly improving the data-locality and chance of a cache-hit on the CPU.
However, a deep-vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).
In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.

\subsection{Previous works and goals of this tudy}
Several recent works, e.g. \cite{CurtisGPU:2017} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC.
The shallow-SIMD and SIMT-vectorization models were investigated in OpenCL and compared to an OpenMP baseline code that was deep-vectorized by simple compiler hints (a.k.a pragmas).
The shallow vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
\textcite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.

@@

\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison.pdf}
    \caption{}
    \label{F:SIMDComp}
  \end{subfigure}
  ~

\end{figure}

\label{results}

>>>>

In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for different states will load sequential memory locations in Fig.~\ref{F:mem} in memory however the $k+1$'th species concentration will be $N$ memory locations away, increasing the likelihood of cache-misses on the CPU.
In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k\dots k+V$ (for a vector width of $V$) will access sequential memory locations if the data is stored in ``C''-order.
Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s$, with $N_s \ll N$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).
In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.

\subsection{Previous works and goals of this tudy}
Several recent works, e.g. \cite{CurtisGPU:2017} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC.
The shallow-SIMD and SIMT-vectorization models were investigated in OpenCL and compared to an OpenMP baseline code that was deep vectorized by simple compiler hints (a.k.a pragmas).
The shallow vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
\textcite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.

@@

\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison}
    \caption{Runtime in milliseconds per thermo-chemical state of shallow-SIMD and SIMT-vectorization on a single CPU core for both ``C'' (filled symbols) and ``F'' (empty symbols) data ordering.}
    \label{F:SIMDComp}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics[width=\textwidth]{SIMD_Vecwidth_comparison}
    \caption{Speedup of a ``C''-ordered shallow-SIMD vectorization on a single CPU core when varying the supplied vector-width.}
    \label{F:Veccomp}
  \end{subfigure} \\
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\textwidth]{SIMD_scaling}
    \caption{Strong parallel scaling of a ``C''-ordered shallow-SIMD vectorization using a vector width of 16.}
    \label{F:SIMDscale}
  \end{subfigure}
\end{figure}

In Fig.~\ref{F:SIMDComp}, the performance of the shallow-SIMD and SIMT vectorization are compared on a single CPU core for both ``C'' and ``F'' memory formats.
The ``C''-ordered SIMD\slash SIMT vectorizations show a speedup over the equivalent ``F''-ordered code by \SIrange{1.36}{1.96}{$\times$} and \SIrange{1.60}{2.38}{$\times$} respectively.
This demonstrates that enhanced data-locality of the ``C''-ordering for CPU caching---discussed in Sec.~\ref{S:data}---is key to high performance on the CPU, and suggests that a deep vectorization would see even greater accelerations.
The shallow vectorization is consistently faster than the SIMT vectorization, by \SIrange{1.72}{1.88}{$\times$} and \SIrange{2.10}{2.42}{$\times$} for the ``C'' and ``F'' memory formats respectively.
This is less than the theoretical maximum speedup on this CPU, which can process \SI{256}{\bit} AVX instructions (i.e. four double-precision operations) concurrently.
Most likely this is due to the non-sequential memory access patterns of the shallow vectorization discussed previously, but could also be caused in part by inefficiencies in translation of the OpenCL code to AVX code, a process handled by the Intel Opencl Runtime.

Figure~\ref{F:Veccomp} investigates the effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core.
Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from memory locations increasingly far apart and the performance of all three vector widths becomes similar.
Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models for a ``C''-ordered shallow vectorization with vector-width of 16:
The scaling efficiency for all models exhibits a power-law dependence, decreasing with the number of cores used.
Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies with isopentanol having the highest efficiency in all cases.
This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
Additionally, OpenCL allows assigning of work to devices by shared resources e.g., a L3 cache, which was not pursued in this work for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
